Human Activity Monitoring Data Analysis
========================================================

First, read the data in using the 'read.csv' function.
```{r}
activity <- read.csv("activity.csv")
```

The tapply function can then be used to summarize the number of steps by date.
```{r}
n <- tapply(activity$steps, activity$date, mean, na.rm = TRUE)
hist(n, main = "Steps Per Day", xlab = "Number of Steps", col = "lavender")
```

The average number of steps per day seem symetrical in the histogram, and indeed, the mean and median are very similar.
```{r}
mean(n, na.rm = TRUE)
median(n, na.rm = TRUE)
```

We could also summarize the steps data by time point instead of by date. Some interesting patterns emerge.
```{r}
m <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
plot(m, type = "l", xlab = "Time (in 5 minute bins)", ylab = "Average Number of Steps")
```

The names function can be used to retrieve the time point with the maximum number of steps from the 'm' object.
```{r}
names(m[m==max(m)])
```

Unfortunately, there is a lot of missing data in this data set.
```{r}
sum(is.na(activity$steps))
```

Using the 'm' object we created previously, we can create a new data frame (called 'nas.imputed') and fill in all NAs with the mean number of steps for that time of day. 

```{r}

nas.imputed <- activity

for (row in 1:nrow(nas.imputed)) {
  time <- as.character(nas.imputed[row, 3])
  if (is.na(nas.imputed[row, 1])) {
      nas.imputed[row, 1] <- m[[time]]
  }
}

head(nas.imputed)
```

As we can see, this method of filling in NAs has little impact on the total number of steps per day.

```{r}
l <- tapply(nas.imputed$steps, nas.imputed$date, mean, na.rm = TRUE)
hist(l, main = "Steps Per Day (Missing Values Imputed)", xlab = "Number of Steps", col = "lavender")
mean(l, na.rm = TRUE)
median(l, na.rm = TRUE)
```

Finally, we can ask whether activity changes between weekends and weekdays. 

```{r}
nas.imputed$date <- as.Date(nas.imputed$date)
nas.imputed$weekend <- ifelse(weekdays(nas.imputed$date) == "Saturday" | weekdays(nas.imputed$date) == "Sunday", "weekend", "weekday")
```

